---
title: "RansomwareAttacksAnalysis"
author: "Zack Espinoza, Christian Garza"
date: "2024-04-16"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load data

```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(dslabs)

#<<<<<<< HEAD
#=======
#install.packages("patchwork") #run if patchwork not downloaded
library(patchwork)
#=======
 #Installed
#>>>>>>> eb922c317e58f6a713f58f0c22e0296e7311011c
RA <- read.csv("Ransomware Attacks - Ransomware Attacks.csv") #Ransomware Attacks
bitcoinHigh <- read.csv("bitcoinHighPrice.csv") #for bitcoin cost graph
```

# Introduction

Ransomware refers to a type of malware that locks a user out of their
system and can hold their data hostage. The expectation is that the
victim will pay a ransom for the attacker to release the data back to
the user or withhold it from the public. With the rise of
cryptocurrency, these attacks have become far more common than the past
due to the unregulated nature of coins so criminals can receive a ransom
and hide their tracks. Our goal is to analyze some of these attacks and
find out some answers about both the attacks, as well as common victims
and outcomes of these attacks.

# Explaining the Data

Due to the nature of these attacks, many of the examples are going to
have unknown data such as ransom amount, this is because companies do
not like making it public that their security has been breached unless
necessary or the attacker has made their intentions public, or worst
case scenario, leaked the data to the public or sold it on the black
market. This could mean that our data does not apply to the whole
situation of ransomware attacks, but we can make conclusions about the
data we do know.

# Questions: Christian(1, 2, 4, 5) Zack(3, 6, 7)

1.  What industries are attacked by Ransomware the most?
2.  Where are these attacks coming from
3.  Does a company's revenue or size have any correlation with the
    ransom amount?
4.  Which regions are being attacked the most.
5.  Which hacker groups are attacking which region most frequently,
    Europe, America, Asia/Oceania.
6.  Average ransom amount vs average amount actually paid.
7.  Is ransomware that we know about trending down or gaining
    popularity? Is this changed by the current price of bitcoin?

# Unfiltered Data

```{r}
glimpse(RA)
```

# Explaining the Columns

-   **Target/AKA/description** - Name of the organization attacked
-   **Sector** - What kind of company they are, healthcare, tech, etc
-   **Size** - Number of employees
-   **revenue** - revenue of the company in millions USD
-   **cost/ransom** - cost of the ransom and explaination in millions
-   **Ransomware** - the individual or group taking credit for the
    attack

### Nullifying irrelevant info from dataset

```{r}

RA$AKA <- NULL
RA$description <- NULL
RA$organisation.size.1.5.10.25.100.300. <- NULL
RA$data.note <- NULL
RA$YEAR.code <- NULL
RA$interesting.story..edited. <- NULL
RA$interesting.story. <- NULL
RA$stock.symbol <- NULL
RA$interesting.story..long. <- NULL
RA$revenue.as.of <- NULL
RA$no.of.employees <- NULL
RA$Data.Note <- NULL
RA$Source.Name <- NULL
RA$URL <- NULL
RA$URL.2 <- NULL
RA$URL.3 <- NULL
RA$URL.4 <- NULL
RA$URL.5 <- NULL
RA$X.ID <- NULL
```

# 1. What Industries Are Attacked By Ransomware The Most?

```{r plot-wider, fig.width=12, fig.height=13}
most_affected <- RA %>% group_by(sector) %>% summarise(n=n()) %>% arrange(desc(n), sector)

ggplot(most_affected, aes(y=fct_rev(sector), x = n)) + geom_col() + scale_x_continuous() + theme(panel.background = element_rect(color = "white", fill = "gray")) + labs(y=NULL, x = "Number of Attacks") + labs(title = "Industries Affected By Ransome Attacks", subtitle = "Ransome Attacks From 2013-2023")
```

(Explain how once we have all the data we can narrow it for what we
want)

# Top 5 Industries Attacked By Ransomware The Most

```{r}
Top_5_RA <- RA %>% group_by(sector) %>% summarise(n=n()) %>% arrange(desc(n)) %>% head(5)

most_affected <- RA %>% group_by(sector) %>% summarise(n=n()) %>% arrange(desc(n), sector) %>% slice(1:5)
               ggplot(combinedData, aes(x = factor(YEAR))) +
  geom_bar(aes(y = ..count..), fill = "black")
               
ggplot(most_affected, aes(x=(n), y = sector, fill = "red")) + geom_col() + scale_x_continuous() + theme(panel.background = element_rect(color = "white", fill = "gray")) + labs(y=NULL, x = "Number of Attacks") + labs(title = "Top 5 Industries Affected By Ransome Attacks", subtitle = "Ransome Attacks From 2013-2023")
```

(Explain)

# 2. Where are these attacks coming from

```{r}
RA_Hackers <- RA %>% group_by(Ransomware) %>% summarise(n=n())

Temp_RA_ATT <- RA_Hackers$region <- c("N/A", "USA",	"Japan",	"Russia",	"Russia",	"Russia",	"Russia",	"Russia",	"Iran",	"Russia",	"Russia",	"Russia",	"Russia",	"Russia",	"Cuba",	"Russia",	"Russia",	"UK",	"Ukraine",	"USA",	"Turky",	"Russia",	"Russia",	"Ukraine",	"Ecuador",	"Russia",	"Russia",	"Russia",	"Russia",	"Russia",	"Russia",	"Russia",	"Russia",	"N/A",	"Russia",	"Russia",	"Netherlands",	"Iran",	"Russia",	"Russia",	"Russia",	"Russia",	"Spain",	"Russia",	"Russia",	"Russia",	"Russia",	"Russia",	"Russia",	"USA",	"Iran",	"South Africa",	"Russia",	"Russia",	"Russia",	"Russia",	"North Korea",	"Russia",	"Chinese",	"Russia",	"N/A")

RA_loc_of_att <- merge(RA, RA_Hackers, by = c("Ransomware"), all.x = TRUE)

RA_loc_of_att$n <- NULL
RA_V.2 <- RA_loc_of_att[, c(2, 3, 8, 9, 10, 1, 11, 6, 7, 4, 5)]
RA_Attacker_Map <- RA_V.2 %>% group_by(region) %>% summarise()


mapdata. <- map_data("world")

mapdata <- left_join(mapdata., RA_Hackers, by="region")
mapdata1 <- mapdata %>% filter(!is.na(mapdata$Ransomware))
map1 <- ggplot(mapdata1, aes(x= long, y = lat, group=group)) +
  geom_polygon(aes(fill = region),fill = "firebrick4", color = "black") + 
  borders(alpha = 0.1, fill = "dodgerblue") +
  theme(panel.background = element_rect(color = "black", fill = "gray")) +
  labs(title = "World Map ", subtitle = "Locations Where Ransome Attacks originate from")

map1
```

```{r}
RA_Attacker_Map #List of Ransome Countries (WIP) Need to make Key
```

# 3. Does a company's revenue or size have any correlation with the ransom amount?

```{r}

scatterPlot <- RA %>% ggplot(aes(x = revenue..USD.million, y = ransom.cost, color = sector, na.rm = T)) +
  geom_point() +
   labs(
    x = "Revenue(Hundred Millions USD)" ,
    y = "Ransom(Millions USD)"
  ) +
  scale_x_continuous(labels = scales::number_format(scale = 1e-3), limits = c(0,200)) +
  scale_y_continuous(limits = c(0, 100))


scatterPlot


```

There does not seem to be any relation between the revenue of a company
and the requested ransom amount, nor does it have any correlation with
what sector the company is in.

# 4. Which regions are being attacked the most.

```{r}
RA_Victums <- RA_V.2 %>% group_by(location, sector) %>% summarise(n=n())
names(RA_Victums)[1] <- "region"


Xmapdata. <- map_data("world")


Xmapdata <- left_join(Xmapdata., RA_Victums, by="region")
Xmapdata1 <- Xmapdata %>% filter(!is.na(Xmapdata$sector))
Xmap1 <- ggplot(Xmapdata1, aes(x= long, y = lat, group=group)) +
  geom_polygon(aes(fill = region),fill = "darkgreen", color = "black") + 
  borders(alpha = 0.1, fill = "dodgerblue") +
  theme(panel.background = element_rect(color = "black", fill = "gray")) +
  labs(title = "World Map ", subtitle = "Ransome Attacks Targets")
Xmap1

#Not Sure about this one
```

(This where they all are)

```{r}
df <- tribble(
  ~Country, ~Value,
  "Argentina",	2,
"Australia",	14,
"Austria",	1,
"Belgium",	2,
"Brazil",	7,
"Canada",	15,
"Chile",	2,
"China",	5,
"Croatia",	1,
"Denmark",	2,
"Ecuador",	2,
"Europe",	1,
"Finland",	2,
"France",	14,
"Germany",	13,
"Greece",	1,
"Hong Kong,",	1,
"Hungary",	1,
"India",	8,
"Indonesia",	1,
"Ireland",	4,
"Israel",	3,
"Italy",	2,
"Japan",	9,
"Malaysia",	2,
"Mexico",	2,
"New Zealand",	1,
"Norway",	4,
"Pakistan",	1,
"Phillipines",	1,
"Poland",	1,
"Portugal",	2,
"Russia",	7,
"Scotland",	3,
"Seychelles",	1,
"Singapore",	1,
"Slovakia",	1,
"South Korea",	4,
"Spain",	6,
"Sweden",	1,
"Switzerland",	2,
"Taiwan",	4,
"Thailand,",	1,
"The Netherlands",	5,
"UK",	29,
"Ukraine",	2,
"USA",	165,
"Worldwide",	5

)
map_data("world") %>% 
 
  left_join(df, by = join_by(region == Country)) %>% 
  ggplot(aes(long, lat)) +
  geom_polygon(aes(group = group, fill = Value), color ="black") +
  scale_fill_gradient(low = "skyblue", 
                      high = "darkblue", 
                      limits = c(1,165), 
                      na.value = "gray90") + 
  coord_sf() +
  theme_bw() +
  labs(title = "World Map ", subtitle = "Ransome Attacks Targets")
```

(This one is good and hows many attacks in each country) Its North
America.

# 5.A Which hacker groups are attacking which region most frequently, Europe, America, Asia/Oceania.

```{r}
RA_Victums <- RA_V.2 %>% group_by(location, sector) %>% summarise(n=n())
names(RA_Victums)[1] <- "location"

RA_Big_Victim <- RA_V.2 %>% group_by(location, Ransomware) %>% summarise()

```



# Average ransom amount vs average amount actually paid.

```{r}

avgRansomCost <- RA %>% summarise(requested = mean(ransom.cost, na.rm = T))
avgRansomPaid <- RA %>% filter(ransom.paid == "ransom paid") %>%  summarise(paid = mean(ransom.cost, na.rm = T))
avgRansomUnpaid <- RA %>% filter(ransom.paid == "refused") %>%  summarise(refused = mean(ransom.cost, na.rm = T))

print("The average of all known ransom requests(in millions USD")
avgRansomCost
print("The average of paid ransoms(in millions USD)")
avgRansomPaid
print("The average of refused ransoms(in millions USD)")
avgRansomUnpaid

```

The paid average is significantly lower than the overall average and the
refused average.

# Is ransomware that we know about trending down or gaining popularity? Is this changed by the current price of bitcoin?

```{r}

yearlyAttacks <- RA %>% filter(YEAR != 2023) %>%  group_by(YEAR) %>% ggplot(aes(x = factor(YEAR))) + 
  geom_bar(fill = "black") +
  labs(
    x = "Year" ,
    y = "Number of Attacks"
  )

bitcoinCost <- bitcoinHigh %>% group_by(YEAR) %>% ggplot(aes(x = factor(YEAR), y = high, group = 1)) + 
  geom_line(color = "gold") +
  labs(
    x = "Year" ,
    y = "Bitcoin Peak(USD)"
  )

yearlyAttacks / bitcoinCost

combinedData <- merge(RA, bitcoinHigh, by = "YEAR")

combinedPlot <- ggplot(combinedData, aes(x = factor(YEAR))) +
  geom_bar(aes(y = ..count..), fill = "black") +  
  geom_line(aes(y = high * .001, group = 1), color = "gold") +  
  scale_y_continuous(
    name = "Attacks",
    sec.axis = sec_axis(~.*1000, name = "Bitcoin Peak(USD)"),
    
  ) +
  labs(
    x = "Year"
  )


combinedPlot

```

The trend of bitcoin and number of attacks actually does seem to be
similar, especially when it dips in 2018 and goes back up in 2020
